Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-04-03 paper/manuscript.Rmd
reflow
fig2 caption
minor edits

The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to
approximately 300 woodland caribou. During routine fieldwork on May 30, 2017, we
observed an unmarked adult male caribou swim between Western and Eastern Indian
Islands (Figure 1), a distance of at least 470 m which took ~9 minutes (~52 m
per minute). This observation prompted us to investigate whether swimming
behaviour was common among GPS radio-collared caribou in the population (n = 29
adult females; 1240g, GPS 4400M; Lotek Wireless Inc., Newmarket, Ontario,
Canada). Using `r substr(version$version.string, 1, 15)` [@R], we identified swimming events as two consecutive GPS locations (2-hour
relocation rates)  from an individual occurred on different islands.
Northeastern Newfoundland typically experiences pack ice during winter and
caribou are known locally to travel between islands by walking over the ice. We
restricted our GPS data to the ice-free period of the year (April 1 - December
31) and we identified `r nids` adult female caribou that collectively had `r nevents` 
swimming events over three years (Figure 2). In addition to our own
visual and remotely sensed observations, residents of Fogo Island have also
reported observing caribou swimming between islands on numerous occasions.

@@

(ref:fig2) A) Map of the Fogo Island archipelago with swimming events between islands. B) Histogram displaying the distribution of swimming events throughout the year. Note, colours correspond to individual caribou. C) Inset of swimming events between small islands on the northern coast of Fogo island. D) Inset of swimming events between islands on the southern coast of Fogo island.

>>>>

The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to
approximately 300 woodland caribou. During routine fieldwork on May 30, 2017, we
observed an unmarked adult male caribou swim between Western and Eastern Indian
Islands (Figure 1, see Figure 2 for location), a distance of at least 470 m which took ~9 minutes (~52 m
per minute). This observation prompted us to investigate whether swimming
behaviour was common among GPS radio-collared caribou in the population (n = 29
adult females; 1240g, GPS 4400M; Lotek Wireless Inc., Newmarket, Ontario,
Canada). Using `r substr(version$version.string, 1, 15)` [@R], we identified
swimming events as two consecutive GPS locations (2-hour relocation rates)  from
an individual occurred on different islands. Northeastern Newfoundland typically
experiences pack ice during winter and caribou are known locally to travel
between islands by walking over the ice. We restricted our GPS data to the
ice-free period of the year (April 1 to December 31) and we identified `r nids`
adult female caribou that collectively had `r nevents` swimming events over
three years (Figure 2). In addition to our own visual and remotely sensed
observations, residents of Fogo Island have also reported observing caribou
swimming between islands on numerous occasions.

@@

(ref:fig2) A) Map of the Fogo Island archipelago with swimming events between islands. B) Histogram displaying the distribution of swimming events throughout the year. Note, colours correspond to individual caribou. C) Inset of swimming events between small islands on the northern coast of Fogo island. D) Inset of swimming events between islands on the southern coast of Fogo island. Note, the grey diamond represents the location we observed an adult male caribou swimming (see Figure 1).
